---
  title: "Frequency of cells per cluster"
author: "Emma Bishop and Brooke"
output:
  html_document:
  toc: yes
toc_depth: 4
toc_float:
  collapsed: no
date: "version `r format(Sys.time(), '%B %d, %Y')`"
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

Differential gene expression (DGE) analysis using pseudo-bulking and DESeq2. This will help inform if any annotated clusters can be collapsed, which would help simplify cell-cell communication analysis and DGE re-analysis.

# Load libraries and data

```{r, message=FALSE}
library(Seurat)
library(tidyverse)
library(CellChat)
library(svglite)
library(ggrepel)
library(gridExtra)
library(ggpubr)

set.seed(4)

script_output_dir <- "C:/Users/bdk45/OneDrive - UW/KennedyBrooke/Projects/BCG challenge (INH & RIF)/Rscript"
```

Load integrated, annotated data

```{r, message=FALSE, warning=FALSE}
bcgskin_integ <- readRDS(file = file.path("C:/Users/bdk45/OneDrive - UW/2022_BCGChallenge/Analysis/bcg_challenge_out_both_gex/processed_data/9_bcgskin_annot.rds"))
```

# Prep data

* Untreated PTIDs
* Enough cells (10+) and unique PTIDs (2+) at each timepoint for each cell type

Modify metadata to make life easier downstream.

```{r}
bcgskin_integ@meta.data <- bcgskin_integ@meta.data %>%
  mutate(
    # Add ARM info form control data so we can subset by that
    ARM = case_when(
      Timepoint == "CTRL" ~ "CTRL",
      .default = ARM),
    # Replace underscores with dashes which will happen anyway during aggregation
    eb_manual = gsub("_", "-", eb_manual),
    # Append some characters so no PTIDs start with a number
    PTID = paste("ptid", PTID, sep = ""),
    # Fix the fact that 27 was mislabelled
    umap_label = case_when(
      umap_legend == "27: Tcell_CREM+" ~ "27",
      .default = umap_label)
  )

bcgskin_integ@meta.data <- bcgskin_integ@meta.data %>%
  mutate(umap_label = factor(umap_label, levels = c("0", "1", "2", "3", "4", "5", 
                                                    "6", "7", "8", "9", "10", "11",
                                                    "12", "13", "14", "15", "16", 
                                                    "17", "18", "19", "20", "21",
                                                    "22", "23", "24", "25", "26", "27"))
  )
```

Subset, keeping just untreated PTIDs. This includes healthy control, day 3, and day 15 PTIDs.

```{r}
untx <- subset(bcgskin_integ, subset = ARM %in% c("CTRL", "Non-INH", "Non-RIF"))
Idents(untx) <- "eb_manual"
```

# UMAP of untreated cells

```{r}
# Just have axis lines in the corner
axis <- ggh4x::guide_axis_truncated(
  trunc_lower = unit(0, "npc"),
  trunc_upper = unit(1.75, "cm")
)

umap_plt <- DimPlot(untx, label = TRUE, 
                    group.by = "umap_label", pt.size = 0.005) +
  NoLegend() +
  guides(x = axis, y = axis) +
  theme(text = element_text(family="Arial"),
        plot.title = element_blank(),
        axis.line = element_line(arrow = grid::arrow(length = unit(0.3, "cm"),
                                                     ends = "last")),
        axis.title = element_text(hjust = 0.01),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        plot.margin = grid::unit(c(0,0,0,0), 'cm')) +
  scale_x_discrete("UMAP1") +
  scale_y_discrete("UMAP2")


umap_legend_plt <- DimPlot(untx, label = TRUE, 
                           group.by = "umap_legend", pt.size = 0.005) +
  guides(x = axis, y = axis) +
  theme(text = element_text(family="Arial"),
        plot.title = element_text(size = 18, face = "plain"),
        axis.line = element_line(arrow = grid::arrow(length = unit(0.3, "cm"),
                                                     ends = "last")),
        axis.title = element_text(hjust = 0.01),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.margin = grid::unit(c(0,0,0,0), 'cm')) +
  scale_x_discrete("UMAP1") +
  scale_y_discrete("UMAP2")

umap_legend <- ggpubr::get_legend(umap_legend_plt)
umap_legend <- as_ggplot(umap_legend)

umap_final <- umap_plt + umap_legend
umap_final
```


# Make table of cell counts per PTID, timepoint, and cluster 
````{r,message=FALSE}
#Combine fibroblast clusters 
untx@meta.data <- untx@meta.data %>%
                  mutate(eb_manual = case_when(umap_label == "6" ~ "Fibroblast", 
                                               umap_label == "7" ~ "Fibroblast",
                                               .default = eb_manual))

cell_counts <- as.data.frame(table(untx$PTID,untx$eb_manual))
colnames(cell_counts) <- c("PTID","Cell_type","Frequency")
cell_counts <- pivot_wider(cell_counts, names_from="Cell_type", values_from = "Frequency")

total_cells <- as.data.frame(table(untx$PTID))
colnames(total_cells) <- c("PTID","Total_cells")

cell_counts <- merge(cell_counts, total_cells, by="PTID")

remove(total_cells)
```

# Transform cell counts into frequencies of total cells per patient 
````{r,message=FALSE}
freq_counts <- cell_counts %>% 
                mutate()

```

# Make boxplot of frequency and calculate P-values 
````{r,message=FALSE}
colnames(Fibroblast_MMP1_frequencies) <- c("PTID","Group","Count","Proportion")

ggplot(Fibroblast_MMP1_frequencies, aes(x = Group, y = Proportion, fill = Group)) +
    geom_boxplot() + 
    scale_fill_brewer(palette = "Set3") +
    labs(title = "Frequency of MMP1+ Fibroblasts",
         x = "Group",
         y = "Proportion of Total Cells (%)") +
    theme_minimal() +
    stat_compare_means(method = "wilcox.test", label.x = 1.5) 

````


# Cassie's code 
```{r}
library(tidyverse) 
library(dplyr) 
library(ggplot2) 
library(cowplot) 
library(viridis) 
library(ggpubr) 
library(here) 
library(RColorBrewer) 
library(rstatix) 
library(Cairo) 

## Read in data ## 
# Background subtracted cytokine specific data 
f50_ifnghi_cd1b <- read_csv(file = "out/dataframes/Frequencies_bgsub_50ngmL_IFNg_hi_cd1b") 

## Merge data, excluding populations except CD1b MA and CD1b GMM specific T cells ## 
f50_ifnghi_tet <- f50_ifnghi_cd1b %>%   filter (Cell_type == "CD1bMA_IFNg_hi" | Cell_type == "CD1bGMM_IFNg_hi") 

## Perform statistical testing ## 
Fibroblast_MMP1_frequencies <- Fibroblast_MMP1_frequencies %>%   
dunn_test(Timepoint ~ Proportion, p.adjust.method = "bonferroni") %>%   
adjust_pvalue() %>%   
add_significance() %>%   
filter(p.adj.signif != "ns") 

## Make plots ## 
# Palettes 
palette2 <- c("lightblue3", "darkolivegreen")

# Specify cell_type order 
f50_ifnghi_tet$Cell_type <- factor(f50_ifnghi_tet$Cell_type, levels = c("CD1bMA_IFNg_hi", "CD1bGMM_IFNg_hi")) 

# Barplot 
ifng_tet <- ggbarplot(f50_ifnghi_tet, x = "Cell_type",           
y = "IFNg_hi_freq",           
color = "black", fill ="Cell_type", palette = palette2,           
add = c("mean_sd", "jitter")) +   
labs(y = "Induced IFN\U03B3 Expression") +   
scale_x_discrete(name = " ", labels=c("CD1bGMM_IFNg_hi" = "CD1b \nGMM",                                         "CD1bMA_IFNg_hi" = "CD1b \nMA")) +   stat_pvalue_manual(f50_ifnghi_tet_s, label = "p.adj.signif", y.position = 99, step.increase = 0.065, size = 7) +   
scale_y_continuous(n.breaks = 6, limits = c(-1,100)) +   
theme_bw(base_size = 24) +   
theme(legend.position = "none") +   
ggtitle("Ex-vivo PBMC") 

#Save plot
png("out/plots/IFNg_hi_tet_bgsub.png", 400, 425) 
plot(ifng_tet) 
dev.off()
```